---
title: "R code for Area Under the Curve (AUC)"
output: html_document
date: "2023-11-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("dplyr")
#install.packages("ggplot2")
library(dplyr)
library(ggplot2)
```

## Function created for finding the area under the curve (AUC)

```{R, function for finding Area under the curve, echo=TRUE}
area_under_curve <- function (x, y) {
  
  n <- length(y)
  
  res <- 0
  
  for (i in 1:(n-1)) {
    
    res <- res + ((x[i+1] - x[i]) * (y[i] + y[i+1]))
    
  }
  
  res <- res / 2
  
  return(res)
}


```

### Another version of the AUC function, same effect written shorter

```{R, function for finding Area under the curve2, echo =TRUE}

area_under_curve2 <- function (x, y) {
  
  id <- order(x)
  res <- sum(diff(x[id])* rollmean(y[id],2))
  
  return(res)
  
}
```

## Example data

Made up dataframe on cows:

```{r data frame}
df <-
tibble(
  condition = c(1, 1, 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 2, 2),
  cow_id = c(1, 1, 1, 1, 2, 2, 2, 2, 1, 1, 1, 1, 2, 2, 2, 2),
  x = c(1, 5, 7, 10, 3, 5, 7, 15, 1, 5, 5.5, 10, 1, 5, 7, 10),
  y = c(2, 6, 3,  8, 2, 3, 3,  2, 5, 3, 3,  8, 2, 4, 3,  1)
) %>%
  group_by(condition, cow_id) %>%
  arrange(x) %>%
  ##an additional colulmn named auc calculated from the function above is now associated with each cow
  mutate(auc = area_under_curve(x, y)) %>%
  ungroup() %>%
  mutate(cow_id = factor(cow_id), condition = factor(condition))
```

## Example plots of the data.

### AUC is quantified and put at the top of each graph in Plot 1 as well as visually demonstrated in the black box

### Plot 2 shows the AUC by the conditions and x, y coordinates

### Plot 3 and 4 show bar graphs of the cow id dodged over condition or visversa

```{r plots, echo=FALSE}

Plot1 <- ggplot(data = df, aes(x = x, y = y)) +
  geom_line() +
  geom_area() +
  facet_wrap(vars(cow_id, auc)) +
  scale_y_continuous(limits = c(0, NA))


Plot2 <- ggplot(data = df, aes(x = x, y = y, colour = cow_id)) +
  geom_line() +
  facet_wrap(vars(condition), labeller = label_both) +
  scale_y_continuous(limits = c(0, NA))
  

Plot3 <- ggplot(data = df, aes(x = condition, y = auc, fill = cow_id, group = cow_id)) +
  geom_col(position = "dodge")


Plot4 <- ggplot(data = df, aes(x = cow_id, y = auc, fill = condition, group = condition)) +
  geom_col(position = "dodge")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
